---
title: "How Americans like their steak"
author: "Peer Christensen"
date: "24/1/2017"
output: github_document
---
```{r,echo=F}
library(fivethirtyeight)
library(RColorBrewer)
```

The data used for this report came from.........

## Steak preparation preference by region

```{r eval=T}
steak_reg=prop.table(table(steak_survey$region,steak_survey$steak_prep),1)
steak_reg=steak_reg[,order=c(5,4,3,2,1)]
par(mfrow=c(3,3),mar=c(3,3,3,3))
cols=brewer.pal(5,"OrRd")
for (i in seq(9)) {barplot(steak_reg[i,],main=dimnames(steak_reg)[[1]][i],col=cols)}
```

## By gender

```{r}
steak_gen=prop.table(table(steak_survey$female,steak_survey$steak_prep),1)
steak_gen=steak_gen[,order=c(5,4,3,2,1)]
par(mfrow=c(1,2))
```
barplot(steak_gen[2,],col=brewer.pal(5,"RdPu"),main="Women")
barplot(steak_gen[1,],col=brewer.pal(5,"Blues"),main="Men")

## Smokers vs. non-smokers

```{r}
steak_smoke=prop.table(table(steak_survey$smoke,steak_survey$steak_prep),1)
steak_smoke=steak_smoke[,order=c(5,4,3,2,1)]
par(mfrow=c(1,2))
barplot(steak_smoke[2,],col=brewer.pal(5,"Greys"),main="Smokers")
barplot(steak_smoke[1,],col=brewer.pal(5,"Greens"),main="Non-smokers")
```
barplot(steak_smoke[2,],col=brewer.pal(5,"Greys"),main="Smokers")
barplot(steak_smoke[1,],col=brewer.pal(5,"Greens"),main="Non-smokers")

## Gamblers vs. non-gamblers
```{r}
steak_gamble=prop.table(table(steak_survey$gamble,steak_survey$steak_prep),1)
steak_gamble=steak_gamble[,order=c(5,4,3,2,1)]
par(mfrow=c(1,2))

```
barplot(steak_gamble[2,],col=brewer.pal(5,"Spectral"),main="Gamblers")
barplot(steak_gamble[1,],col=brewer.pal(5,"PiYG"),main="Non-gamblers")

## Drinkers vs. non-drinkers

```{r}
steak_alcohol=prop.table(table(steak_survey$alcohol,steak_survey$steak_prep),1)
steak_alcohol=steak_alcohol[,order=c(5,4,3,2,1)]
par(mfrow=c(1,2))
```
barplot(steak_alcohol[2,],col=brewer.pal(5,"BuPu"),main="Drinkers")
barplot(steak_alcohol[1,],col=brewer.pal(5,"PuRd"),main="Non-drinkers")

## By age

```{r}
steak_age=prop.table(table(steak_survey$age,steak_survey$steak_prep),1)
steak_age=steak_age[,order=c(5,4,3,2,1)]
par(mfrow=c(1,4),mar=c(3,3,3,3))
cols=brewer.pal(5,"YlGn")
```
for (i in seq(4)) {barplot(steak_age[i,],main=dimnames(steak_age)[[1]][i],col=cols)}

## By household income group

```{r}
steak_hhold_income=prop.table(table(steak_survey$hhold_income,steak_survey$steak_prep),1)
steak_hhold_income=steak_hhold_income[,order=c(5,4,3,2,1)]
par(mfrow=c(1,5),mar=c(3,3,3,3))
cols=brewer.pal(5,"Reds")
```
for (i in seq(5)) {barplot(steak_hhold_income[i,],main=dimnames(steak_hhold_income)[[1]][i],col=cols)}

## By level of education

```{r}
steak_educ=prop.table(table(steak_survey$educ,steak_survey$steak_prep),1)
steak_educ=steak_educ[,order=c(5,4,3,2,1)]
par(mfrow=c(1,4),mar=c(3,3,3,3))
cols=brewer.pal(5,"GnBu")
#drop less than High school level - only one respondent
```
for (i in seq(2,5)) {barplot(steak_educ[i,],main=dimnames(steak_educ)[[1]][i],col=cols)}

## Separating smokers, gamblers and drinkers from the rest

For this visualization, we create a new variable to distinguish people
who enjoy smoking, gambling and drinking, from those who do not like fun. All respondents who do not fit into these categories are disregarded.

```{r}
steak_survey$champs=with(steak_survey,
                         ifelse(gamble==T & smoke==T & alcohol==T,T,NA))
steak_survey$champs[steak_survey$gamble==F & 
                steak_survey$smoke==F & steak_survey$alcohol==F] = F
steak_champs=prop.table(table(steak_survey$champs,steak_survey$steak_prep),1)
steak_champs=steak_champs[,order=c(5,4,3,2,1)]
par(mfrow=c(1,2))
```
barplot(steak_champs[2,],col=brewer.pal(5,"YlOrBr"),main="Champs")
barplot(steak_champs[1,],col=brewer.pal(5,"Purples"),main="Non-champs")
